### 2011-08-15 src/article.tex
Revise Background

% BioMed_Central_Tex_Template_v1.05 %{{{1

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmc_article}

% Load packages
\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage{trackchanges}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}

\def\includegraphic{}
\def\includegraphics{}

\setlength{\topmargin}{0.0cm}
\setlength{\textheight}{21.5cm}
\setlength{\oddsidemargin}{0cm} 
\setlength{\textwidth}{16.5cm}
\setlength{\columnsep}{0.6cm}

\newboolean{publ}

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

% Scaffolder url
\urldef{\scaffolder}\url{http://next.gs}

% Begin ...
\begin{document}
\begin{bmcformat}

\title{Scaffolder - Software for \change{Reproducible}{Manual} Genome
Scaffolding.}
%{{{1

\author{
  Michael D Barton$^{1}$\correspondingauthor%
  \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
\and
  Hazel A Barton$^1$%
  \email{Hazel A Barton - bartonh@nku.edu}%
      }

\address{\iid(1) Department of Biological Sciences, Northern Kentucky %
University, Nunn Drive, Highland Heights, KY 41076 }%

\maketitle

\clearpage

\begin{abstract} %{{{1

  \paragraph*{Background:} \add{The} assembly of \add{next-generation}
  short-read sequencing data can result in a fragmented non-contiguous
  \change{series}{set} of genomic sequences. Therefore a common step in
  a genome project is to join neighbouring sequence regions together and fill
  gaps \remove{in the assembly using additional sequences}. This scaffolding
  step\remove{, however,} is non-trivial and requires manually editing large
  blocks of nucleotide sequence. Joining these sequences together also hides
  the source of each region in the final genome sequence. Taken together,
  these considerations may make reproducing or editing an existing genome
  \change{build}{scaffold} difficult.

  \paragraph*{Methods:} The software outlined here, ``Scaffolder,'' is
  implemented in the Ruby programming language and can be installed via the
  RubyGems software management system. Genome scaffolds are defined using YAML
  - a data format\remove{,} which is both human and machine-readable. Command
  line binaries and extensive documentation are available.

  \paragraph*{Results:} This software allows a genome \change{scaffold}{build}
  to be defined in terms of the constituent sequences using a relatively simple
  syntax\remove{to define the scaffold}. This syntax further allows unknown
  regions to be \change{defined}{specified} and \remove{add} additional
  sequence to \add{be used} fill \add{known} gaps in the scaffold. Defining
  the genome construction in a file makes the scaffolding process reproducible
  and easier to edit compared with \add{large} FASTA nucleotide sequences.

  \paragraph*{Conclusions:} Scaffolder is easy-to-use genome scaffolding
  software\change{. This tool}{ which} promotes reproducibility and continuous
  development in a genome project. Scaffolder can be found at \scaffolder.

\end{abstract}

\ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}

\clearpage

\section*{Background} %{{{1
%{{{2

High-throughput sequencing can produce hundreds of thousands to millions of
sequence reads from a genome. At the time of writing, high-throughput
sequencing is limited to producing reads less than 1,000 nucleotides in length.
Therefore to resolve a sequence longer than this, such as a complete genome,
these numerous smaller fragments must be pieced together. The process of
joining reads into longer sequences is the `assembly' stage of a genome project
\cite{miller2010}. \pb

Assembly software takes the nucleotide reads produced by sequencing hardware
and, in the ideal case, outputs a single complete genome sequence composed of
these individual fragments. An analogy for this process is a jigsaw puzzle:
each nucleotide read represents a single piece, and the final genome sequence
is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
\remove{within the genome, }or biased and incomplete sequencing data, may
prevent the genome being assembled into a continuous sequence. This may be due
to \change{not enough}{insufficient}, or multiple different overlaps between
reads and is analogous to missing pieces in the jigsaw or pieces that fit to
multiple other pieces. \pb

The advent of high\add{-}throughput sequencing methods has led to a renewed
interest in algorithms to solve the problem of genome assembly
\cite{pop2008,pop2009}. \remove{Nonetheless, due to }The complexity of merging
large numbers of overlapping reads \add{can lead to} genome assembly software
\remove{may} being unable to produce a complete sequence. Instead, the
algorithm may generate several large assembled regions of sequence (`contigs')
composed of the many individual reads\remove{ that represent assembly into the
longest possible sequences, given the data. Nonetheless, }These contigs
represent a fragmented picture of the genome and \add{therefore} require
additional work to join together into a complete sequence. \pb

The process of \change{ameliorating a draft sequence into a finished continuous
sequence}{finishing a genome sequence} can be expensive in terms of time and
laboratory effort. \change{although}{In some cases} the genomic data present in
a set of generated contigs may be sufficient for many research questions
\cite{branscomb2002}. Nevertheless, a continuous high-quality `finished' genome
sequence does provide a greater depth of information, such as complete
resolution of repeat regions and precise estimates of distances between genomic
elements \cite{parkhill2002,fraser2002}. The process of joining these contigs
together to form a continuous genome sequence is called the `scaffolding' or
`finishing' stage and is the focus of the software described in this article.
\pb

\subsection*{Scaffolding} %{{{2

Scaffolding is the process of joining a series of disconnected contigs into
a complete continuous genome sequence. Due to genomic complexity and missing
data, scaffolding may not ultimately produce a final completed sequence, but
may still succeed in joining a subset of contigs together or resolving gaps
between contigs. An overview of the required steps in the scaffolding process
is outlined below:

\subsubsection*{Contig Orientation} %{{{3

The sequencing process generates reads from either strand of the DNA helix and
the resulting contigs constructed from these reads may represent either the
5'~$\rightarrow$~3' or 3'~$\rightarrow$~5' DNA strands with respect to the
origin of replication. Orientating all contigs to point in the same direction
requires reverse complementing sequences where necessary. In the case of
\change{microbial}{archaeal and bacterial} genomes this orientation will be to
the 5'~$\rightarrow$~3' direction following the direction of genome
replication.

\subsubsection*{Contig Ordering} %{{{3

Contig ordering determines the placement of observed contigs to best represent
their order in the true genome sequence. The correct placement of each contig
also highlights any extra-genomic DNA, such as plasmids which are scaffolded
separately from the genomic sequence. The order is commonly started at the
contig containing the origin of replication. All subsequent contigs are then
ordered in the 5' $\rightarrow$ 3' direction of DNA replication.

\subsubsection*{Contig Distancing} %{{{3

Given the correct order and orientation, determining the distance between
contigs results in an estimate of the complete genome size. The size of any
inter-contig gaps represents the length of an unknown region in the genome.
Filling these regions with unknown nucleotide characters (`N') allows a draft
continuous sequence. \change{which}{This sequence} is useful for representing
both the known and to-be-resolved areas in the genome sequence.

\subsubsection*{Gap Closing} %{{{3

During the scaffolding process, closing and filling gaps between contigs
completes and improves the genome scaffold. Closing gaps may require returning
to the laboratory to perform additional sequencing or using computational
methods of estimating the unknown sequence. This additional sequence is
\remove{then }used to replace the gap between two contigs, joining them into
a single sequence. Once all contigs have been joined and gaps in a scaffold
closed, the genome may be considered finished.

\subsection*{Computational Methods for Scaffolding} %{{{2

The process of \change{resolving a genome sequence and building a}{finishing
a genome} scaffold \remove{may} uses wet laboratory methods, \emph{in silico}
methods, or a combination of both. An example of a computational method might
use the \remove{existing} paired-read data from the sequencing stage.
\remove{where reads were produced in pairs at a known distance apart in the
genome.} The occurrence of paired reads in separate contigs can be used to
\change{probabilistically estimate}{estimate probabilistically} the order and
distance between these contigs. Alternatively, laboratory methods may use PCR
to amplify the unknown DNA in a gap region then use \remove{traditional}
Sanger sequencing to determine the sequence of this gap. Computational methods
are \remove{more} preferable as they are less costly in laboratory time and
materials compared to manual gap resolution, and use the available data that
exists following sequencing \cite{nagarajan2010}. Finally when the scaffold
cannot be completely resolved, \emph{in silico} software packages exist to
suggest the likely primers necessary for PCR amplifying the sequence in gap
regions \cite{gordon2001}. \pb

Examples of \emph{in silico} methods include \remove{using synteny to}
comparing the assembled contigs to a complete reference genome sequence
\change{by}{to} search\remove{ing} for areas of sequence similarity between the
two. Any areas of corresponding sequence with the reference genome can be used
to infer contig placement and build the contigs into a scaffold
\cite{richter2007,zhao2008,assefa2009}. \add{Genomic} recombination
\remove{within each of the two compared genomes} can however reduce the
\change{efficiency}{efficacy} of this. \pb

Repeat regions may also be responsible for multiple gaps when building
a genome sequence; tandemly repeated nucleotide regions in the genome produce
multiple reads with similar sequence. As many assembly algorithms rely on
sequence overlaps between reads to build a contig, the similarity between
repeat-region reads can result in the assembly collapsing into an artificially
short sequence or being ignored by more conservative assembly algorithms. Such
regions can be resolved by using algorithms \add{that} specifically reassemble
the collapsed repeat region \remove{correctly} \cite{mulyukov2002,koren2010}.
\change{while a}{A} related approach uses unassembled sequence reads matching
the regions around a scaffold gap to construct a uniquely overlapping set of
reads across \change{it}{the gap}. \cite{tsai2010}. \pb

Paired-read data can provide an extra level of information about how contigs
may be scaffolded together. Heuristic scaffolding algorithms take advantage of
this data to search for the optimal configuration of contigs in the scaffold
that matches these paired-read distances \cite{dayarian2010,boetzer2011}.
Synteny data from a reference genome can also be combined with this
paired-read data to estimate the best contig configuration \cite{pop2004}. \pb

These described \emph{in silico} methods provide a wide array of approaches for
merging contigs into a larger, continuous scaffold sequence.
\remove{Nevertheless} The scaffolding process may still require manually
inserting additional sequences or further joining contigs using PCR-derived
sequence. Moving and editing large blocks of nucleotide text by hand however
\remove{possibly} introduces human error and precludes any reproducibility of
the steps in this process. \pb

The software outlined here, ``Scaffolder,'' aims to address these problems of
reproducibility by creating a file syntax and software framework for editing
a genome scaffold. Scaffolder uses a specific file format to define how
contigs are joined, additional sequences are inserted, and for the
specification of unknown regions. This syntax allows a scaffold to be updated
by simply editing the scaffold file. As such, scaffolder facilitates
a reproducible \change{scaffolding}{finishing} process and provides a concise
overview of how the final genomic scaffold was constructed. \pb

\clearpage

\section*{Implementation} %{{{1

\subsection*{Code and Dependencies} %{{{2

Scaffolder is written in the Ruby programming language \change{using version
1.8.7}{and tested against versions 1.8.7 and 1.9.2} \cite{ruby-lang}. The
scaffolder package is split into two libraries. The first called ``scaffolder''
which provides the core Scaffolder application programming interface (API). The
second library ``scaffolder-tools'' provides the Scaffolder command line
interface (CLI). \pb

Unit tests were implemented to maintain individual elements of the source code
during development and were written using the Shoulda and RSpec \cite{rspec}
libraries. Integration tests were written to test the Scaffolder software
interface as a whole and were written using the Cucumber library \cite{rspec}.
\pb

The Scaffolder source code is documented using the Yard library \cite{yard}.
Unix manual pages for the command line were generated using the Ronn library
\cite{ronn}. The manipulation of biological sequences in Scaffolder uses the
BioRuby library \cite{goto2010}. A full list of the software dependencies in
Scaffolder can be found in the Gemfile in the root of each source code
directory. \pb

\subsection*{Scaffold File Syntax} %{{{2

The choice of nucleotide sequences comprising the scaffold is specified using
the YAML syntax \cite{yaml}. YAML is a language format using whitespace and
indentation to produce a machine readable structure. As YAML is a standardised
data format, third-party developers have the option to generate a genome
scaffold using any programming language for which a YAML library exists. The
YAML website lists current parsers for languages including C/C++, Ruby, Python,
Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being widely
supported, YAML-formatted scaffold files can be validated for correct syntax
using third-party tools such as Kwalify \cite{kwalify}. \pb

Initial sequencing data assembly may result in an incomplete genome build.
Generation of additional sequences either through PCR or computational methods
means that genome scaffolding may be an on-going process. The scaffold file
should therefore be simple to update manually in addition to being
computationally tractable. This requirement was also best suited to YAML
syntax which is human-readable and simple to edit in a standard text editor.
\pb

The scaffold file takes the form of a list of entries. Each entry corresponds
to a region of sequence used in the final scaffold sequence. Each entry in the
scaffold file may have attributes that define whether a sub-sequence or the
reverse complement of the sequence should be used. The types of attributes
available, and an example scaffold file are outlined in the Results section.
\pb

\subsection*{Input sequence data and scaffolding process} %{{{2

The input data for Scaffolder is nucleotide sequence in FASTA format file.
These nucleotide sequences can be of any length and for example may be
individual sequences, assembled contigs or contigs which have been joined to
into larger scaffolds. The case in which Scaffolder may be most useful is
using the contigs and scaffolded contigs produced from the previously
assembled set of sequencing reads, combined with additional gap filling
sequences produced by PCR or \emph{in silico} methods as outlined in the
Background. \pb

\remove{The scaffolder algorithm reads through the scaffolder file parsing.
Each entry is then converted to corresponding nucleotide sequence by fetching
from the FASTA file and then performing any designated sequence trimming or
reverse complementing. Each of the sequence entries are then joined into
a continuous single super-sequence and returned to the console output.}

\clearpage

\section*{Results} %{{{1

\subsection*{Scaffolder Simplifies Genome Finishing} %{{{2

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in \remove{the} YAML \remove{syntax} specify how these
sequences should be joined\change{. and the}{The} scaffolder software is used
to generate the scaffold sequence from these instructions. In addition to
specifying which contigs are required, the scaffold file allows the contigs to
be edited into smaller sub-sequences or reverse complemented if necessary.\pb

The process of genome finishing may involve producing additional
oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
\remove{and} The Scaffolder \change{allows}{format provides for using} these
additional insert sequences to \remove{be used} to fill \remove{these} gaps.
These inserts can also be treated in the same manner as larger contig
sequences: trimmed and/or reverse complemented to match the corresponding gap
region size and orientation. \pb

The distances between contigs may be estimated from paired-read data or from
mapping the contigs to a reference genome. These inter-contig gap regions are
useful to join separate sequences together by \change{an}{the} estimated
\change{size and }{distance.} The scaffold file allows for the specification
of such unresolved regions by inserting regions of `N' nucleotides into the
scaffold. The use of these regions in the scaffold indicates the unresolved
regions in the build and their approximate size. \pb

\change{The scaffold file specifies how sequence regions are joined together;
however, the sequences themselves are no  stored in the scaffold file and are
instead}{The nucleotide sequences used in the scaffold are} maintained as
a separate FASTA file. \remove{Nucleotide sequences in a FASTA-format file are
the standard output for a genome assembler, and these may be contigs, sets of
contigs scaffolded into larger sequences, or both. }The nucleotide sequences
\remove{stored in the FASTA file} are referenced in the scaffold using the
first word from in the FASTA header of the corresponding sequence. Maintaining
the nucleotide sequences in a separate file preserves the unedited sequence
and decouples the data from the specification of how it should be used to
produce the genome sequence. \pb

\subsection*{Defining a Scaffold as a Text File} %{{{2

The scaffold file is written using the YAML syntax and an example is shown in
Figure 1. This file illustrates the text attributes used to describe
a scaffold and how the sequences are correspondingly joined together in the
genome build. The basic layout of the scaffold file is a list of entries,
where each entry corresponds to a region of sequence in the generated scaffold
super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
sequence and separate scaffolds should be defined in separate files.} \pb

\subsubsection*{Simple sequence region} %{{{3

The first line of the scaffold file in Figure 1 begins with three dashes to
indicate the start of a YAML-formatted document. The first entry (highlighted
in green) begins with a dash character `-' to denote an entry in the YAML
list. This is a requirement of the YAML format: that each entry begins with
a dash line. The next line is indented by two spaces where whitespace is
required to group similar attributes together. The ``sequence'' tag indicates
that this entry corresponds to a sequence in the FASTA file and the following
line indicates the name of this sequence using the ``source'' tag. The first
word of the FASTA header is used to identify which sequence is selected from
the file. Together these three lines describe the first entry in the scaffold
as a simple sequence using the sequence named `sequence1'. On the right hand
side of Figure 1 this produces the first region in the scaffold, also shown in green. \pb

\subsubsection*{Unresolved sequence region} %{{{3

The second entry in the scaffold, highlighted in orange, is identified by the
``unresolved'' tag, indicating a region of unknown sequence but known length.
The second line specifies the size of this unknown region. This entry produces
a region of 20 `N' characters in the scaffold. \pb

\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags
trim the sequence to these coordinates inclusively.\remove{while} The
``reverse'' tag instructs scaffolder to also reverse complement the sequence.
In the putative scaffold shown in Figure 1 this completes the \remove{scaffold
} sequence. \pb

The final entry in the scaffold \remove{also} includes the ``inserts'' tag to
add additional regions of sequence. These inserts are also added as a YAML
list, with each insert beginning with a dash. The first insert, shown in
purple, uses similar attributes to that of a sequence entry; the reverse,
start and stop tags are used to trim and reverse complement the insert.
Similarly the `source' tag identifies the corresponding FASTA sequence as
`insert1'. The ``open'' and ``close'' tags are specific to inserts and
determine where the insert is added in the enclosing sequence. The region of
the sequence inside these coordinates is inclusively replaced by the specified
insert sequence. This is visualised in the putative scaffold in Figure 1 by
the black lines bisecting the blue sequence. \pb

The next insert, shown in brown, is specified using only the `open' tag. This
illustrates that only one of either `open' or `close' tags is required when
adding an insert sequence. If only one of the `open' or `close' tags is used
the corresponding opposing `open'/`close' coordinate is calculated from the
length of the insert FASTA sequence. This allows inserts to bridge into, and
partially fill gap regions, without requiring an end coordinate position. \pb

\subsection*{Scaffolder Software Interface} %{{{2

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) to validate the
scaffold file and build the draft super sequence. The Scaffolder CLI behaves
as a standard Unix tool and returns appropriate exit codes and manual pages.
The use of both these Scaffolder interfaces is outlined in detail on the
scaffolder website (\scaffolder). \add{This website provides a ``getting
started'' guide as an introduction to using Scaffolder to build a genome
scaffold.} \pb

\clearpage

\section*{Discussion} %{{{1

Scaffolding an incomplete genome assembly requires joining contigs and
additional gap-filling sequence\add{s} using a combination of computational and
laboratory methods. The process of manually editing a scaffold is inherently
hard to reproduce and introduces \remove{the possibility of }irreducible
\change{techniques}{edits} and/or human error. In respect to \change{these
drawbacks}{this} the aims of the Scaffolder software are twofold: 1) to provide
software that is easy to install and simplifies the task of genome finishing;
and 2) to facilitate reproducibility in the scaffolding and finishing stage of
a genome project. \pb

\subsection*{Comparison with other scaffold formats} %{{{2

\remove{Scaffolder was designed to be as simple to use and, assuming the Ruby
and RubyGems software are present, can be installed in a single command line
step. This should negate any possible barriers to entry which require manual
compilation of source code.} Scaffolder \remove{also} uses a minimal and
compact syntax to describe how \change{the}{a} genome scaffold sequence should
be generated. This syntax is simple to \change{construct}{write} and edit
whilst being succinct and readable. \remove{The goal of this syntax is to make
genome scaffolds easy to write in a common text editor.}

\add{AGP is a similar format for describing scaffolds. This format can be used
to define contig order and N-filled gap regions in a scaffold. The advantages
of the AGP format is that each contig entry is defined on a single line which
allows searching the scaffold using UNIX line-based tools. The scaffolder
format in comparison in contrast is written in the standardised YAML format
and therefore accessible to the many languages which provide parsers to this
format. The scaffolder format is provided with a tool explicitly to produce
the fasta sequence of scaffold specified by in the file.} \pb

Constructing a genome using by specifying the scaffold organisation in text
file makes generating a scaffold super sequence both reproducible and
deterministic for the same file and set of FASTA sequences. In comparison,
joining large nucleotide sequences by hand cannot be reliably reproduced. The
scaffold file furthermore provides a human readable description of how the
scaffold is constructed. Configuring the final sequence in the scaffold file
means the build is easier to edit once already constructed. \pb

\subsection*{Example Use Case} %{{{2

\add{The example use case for scaffolder is a combination of computation and
manual editing of a genome scaffold. We have used scaffolder in our genome
projects to create a scaffold file from parsing the output of \emph{in silico}
scaffolding tools into the scaffold file format. This scaffold was then
manually updated in a text editor to fill gap regions with laboratory-produced
sequences.} \pb

The YAML text format \add{also} allows comparison of differences between
scaffold builds using standard Unix tools such as diff. This \add{therefore}
makes scaffold files \add{amenable} to storage in version control systems and
allows genome finishers to use methods similar to those in software
development. \pb

\clearpage

\section*{Conclusions} %{{{1

Scaffolder is software aimed at both bioinformaticians and biologists familiar
with the command line who wish to build a genome scaffold from a set of
contigs. The Scaffolder file format maintains the genome scaffold as a concise
and readable text representation that allows third parties to see how the
genome sequence was scaffolded. This file format also allows a broad overview
of which sequences were included and how they are ordered in the genome
scaffold, something not possible to deduce from a megabase-length string of
nucleotide characters. Scaffolder furthers increases the ease of
reproducibility in genome projects by allowing the scaffold super-sequence to
be reliably reproduced from the same scaffold file. The YAML syntax for
writing the scaffold file is also standardised and simple to manipulate
programmatically. This thereby means the scaffolding process follows the Unix
tenet of ``If your data structures are good enough, the algorithm to
manipulate them should be trivial.''

\clearpage

\section*{Availability and Requirements} %{{{1

  \begin{description}
    \item[Project name:] Scaffolder v0.4.\change{1}{4}, Scaffolder Tools
    v0.1.\change{2}{3}
    \item[Project home page:] \scaffolder
    \item[Operating system:] Platform Independent. Tested on Mac OS X and
    Ubuntu.
    \item[Programming language:] Ruby \add{1.8.7 or 1.9.2}
    \item[Other requirements:] \change{RubyGems}{RubyGems package management
    software and the following libraries: BioRuby 1.4.x, confligliere
    0.1.x, ronn 0.7.x. A full list of development dependencies can be found in
    the Gemfile in the base directory of each project.}
    \item[License:] MIT \item[Any restrictions to use by non-academics:] None
    \end{description}

\clearpage

\section*{Competing Interests} %{{{1

The authors declare no competing interests.

\section*{List of Abbreviations Used} %{{{1

  \begin{description}
    \item[API:] Application programming interface
    \item[CLI:] Command line interface
    \item[PCR:] Polymerase chain reaction
    \item[YAML:] YAML ain't markup language\cite{yaml}
  \end{description}

\section*{Authors contributions} %{{{1

MDB developed and maintains the Scaffolder tool. MDB and HAB wrote the
manuscript.

\section*{Acknowledgements} %{{{1
  \ifthenelse{\boolean{publ}}{\small}{}

This work was supported by the National Institute for Health: IDeA Network of
Biomedical Research Excellence (KY-INBRE) grant (NIH 2P20 RR016481-09) and the
NIH R15 AREA Program grant (R15GM079775).

\clearpage

% Bibliography {{{1
{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{article} }     % Bibliography file (usually '*.bib' )

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

\clearpage

\section*{Figures} %{{{1

\subsection*{Figure 1 - Example of Scaffolder File and the Resulting Build}

An example scaffold file written using the YAML syntax \cite{yaml} (left) and
the resulting putative scaffold sequence (right). The scaffold contains three
entries and two inserts. Each entry in the scaffold file text is delimited by
a `-' on a new line and highlighted using separate colours. The scaffold
diagram on the right is not to scale and instead illustrates how the scaffold
sequences are joined. \pb

\end{bmcformat}

\end{document}