### 2011-09-12 src/article.tex
Spell check

% BioMed_Central_Tex_Template_v1.05 %{{{1

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmc_article}

% Load packages
\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[finalnew]{trackchanges}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}

\def\includegraphic{}
\def\includegraphics{}

\setlength{\topmargin}{0.0cm}
\setlength{\textheight}{21.5cm}
\setlength{\oddsidemargin}{0cm} 
\setlength{\textwidth}{16.5cm}
\setlength{\columnsep}{0.6cm}

\newboolean{publ}

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

% Scaffolder url
\urldef{\scaffolder}\url{http://next.gs}

% Begin ...
\begin{document}
\begin{bmcformat}

\title{Scaffolder - Software for Manual Genome
Scaffolding.}
%{{{1

\author{
  Michael D Barton$^{1,2}$\correspondingauthor%
  \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
\and
  Hazel A Barton$^{1,2}$%
  \email{Hazel A Barton - bartonh@nku.edu}%
      }

\address{
\iid(1) Biology Department, The University of Akron, Akron, %
OH 44325-3908, USA \\
\iid(2) Department of Biological Sciences, Northern Kentucky %
University, Nunn Drive, Highland Heights, KY 41076, USA}

\maketitle

\clearpage

\begin{abstract} %{{{1

  \paragraph*{Background:} The assembly of next-generation short-read
  sequencing data can result in a fragmented non-contiguous set of genomic
  sequences. Therefore a common step in a genome project is to join
  neighbouring sequence regions together and fill gaps . This scaffolding step
  is non-trivial and requires manually editing large blocks of nucleotide
  sequence. Joining these sequences together also hides the source of each
  region in the final genome sequence. Taken together these considerations
  may make reproducing or editing an existing genome scaffold difficult.

  \paragraph*{Methods:} The software outlined here, ``Scaffolder,'' is
  implemented in the Ruby programming language and can be installed via the
  RubyGems software management system. Genome scaffolds are defined using
  YAML - a data format which is both human and machine-readable. Command line
  binaries and extensive documentation are available.

  \paragraph*{Results:} This software allows a genome build to be defined in
  terms of the constituent sequences using a relatively simple syntax. This
  syntax further allows unknown regions to be specified and additional sequence
  to be used to fill known gaps in the scaffold. Defining the genome construction
  in a file makes the scaffolding process reproducible and easier to edit
  compared with large FASTA nucleotide sequences.

  \paragraph*{Conclusions:} Scaffolder is easy-to-use genome scaffolding
  software which promotes reproducibility and continuous development in a
  genome project. Scaffolder can be found at \scaffolder.

\end{abstract}

\ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}

\clearpage

\section*{Background} %{{{1
%{{{2

High-throughput sequencing can produce hundreds of thousands to millions of
sequence reads from a genome. At the time of writing, high-throughput
sequencing is limited to producing reads less than 1,000 nucleotides in
length. Therefore to resolve a sequence longer than this, such as a complete
genome, these numerous smaller fragments must be pieced together. The process
of joining reads into longer sequences is the `assembly' stage of a genome
project \cite{miller2010}. \pb

Assembly software takes the nucleotide reads produced by sequencing hardware
and, in the ideal case, outputs a single complete genome sequence composed of
these individual fragments. An analogy for this process is a jigsaw puzzle:
each nucleotide read represents a single piece, and the final genome sequence
is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
or biased and incomplete sequencing data, may prevent the genome being
assembled into a continuous sequence. This may be due to insufficient, or
multiple different overlaps between reads and is analogous to missing pieces
in the jigsaw or pieces that fit to multiple other pieces. \pb

The advent of high-throughput sequencing methods has led to a renewed interest
in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
The complexity of merging large numbers of overlapping reads can lead to
genome assembly software being unable to produce a complete sequence. Instead,
the algorithm may generate several large assembled regions of sequence
(`contigs') composed of the many individual reads. These contigs represent a
fragmented picture of the genome and therefore require additional work to join
together into a complete sequence. \pb

The process of finishing a genome sequence can be expensive in terms of time
and laboratory effort. In some cases the genomic data present in a set of
generated contigs may be sufficient for many research questions
\cite{branscomb2002}. Nevertheless, a continuous high-quality `finished'
genome sequence does provide a greater depth of information, such as complete
resolution of repeat regions and precise estimates of distances between
genomic elements \cite{parkhill2002,fraser2002}. The process of joining these
contigs together to form a continuous genome sequence is called the
`scaffolding' or `finishing' stage and is the focus of the software described
in this article. \pb

\subsection*{Scaffolding} %{{{2

Scaffolding is the process of joining a series of disconnected contigs into a
complete continuous genome sequence. Due to genomic complexity and missing
data, scaffolding may not ultimately produce a final completed sequence, but
may still succeed in joining a subset of contigs together or resolving gaps
between contigs. An overview of the required steps in the scaffolding process
is outlined below:

\subsubsection*{Contig Orientation} %{{{3

The sequencing process generates reads from either strand of the DNA helix and
the resulting contigs constructed from these reads may represent either DNA
strand. Orientating all contigs to point in the same direction requires reverse
complementing sequences where necessary. In the case of archaeal and bacterial
genomes this orientation will be to the 5'~$\rightarrow$~3' direction following
the direction of genome replication.

\subsubsection*{Contig Ordering} %{{{3

Contig ordering determines the placement of observed contigs to best represent
their order in the true genome sequence. The correct placement of each contig
also highlights any extra-genomic DNA, such as plasmids which are scaffolded
separately from the genomic sequence. The order is commonly started at the
contig containing the origin of replication. All subsequent contigs are then
ordered in the 5' $\rightarrow$ 3' direction of DNA replication.

\subsubsection*{Contig Distancing} %{{{3

Given the correct order and orientation, determining the distance between
contigs results in an estimate of the complete genome size. The size of any
inter-contig gaps represents the length of an unknown region in the genome.
Filling these regions with unknown nucleotide characters (`N') allows a draft
continuous sequence. This sequence is useful for representing both the known
and to-be-resolved areas in the genome sequence.

\subsubsection*{Gap Closing} %{{{3

During the scaffolding process, closing and filling gaps between contigs
completes and improves the genome scaffold. Closing gaps may require returning
to the laboratory to perform additional sequencing or using computational
methods to estimate the unknown sequence. This additional sequence is used to
replace the gap between two contigs, joining them into a single sequence. Once
all contigs have been joined and gaps in a scaffold closed, the genome may be
considered finished.

\subsection*{Computational Methods for Scaffolding} %{{{2

The process of finishing a genome scaffold uses wet laboratory methods,
\emph{in silico} methods, or a combination of both. An example of a
computational method might use the paired-read data from the sequencing stage.
The occurrence of paired reads in separate contigs can be used to estimate
probabilistically the order and distance between these contigs. Alternatively,
laboratory methods may use PCR to amplify the unknown DNA in a gap region then
use Sanger sequencing to determine the sequence of this gap. Computational
methods, using available sequencing data, are preferable as they are less
costly in laboratory time and materials compared to manual gap resolution
\cite{nagarajan2010}. Finally when the scaffold cannot be completely resolved,
\emph{in silico} software packages exist to suggest the likely primers
necessary for PCR amplifying the sequence in gap regions \cite{gordon2001}. \pb

Examples of \emph{in silico} methods include comparing the assembled contigs
to a complete reference genome sequence to search for areas of sequence
similarity between the two. Any areas of corresponding sequence in the
reference genome can be used to infer contig placement and build the contigs
into a scaffold \cite{richter2007,zhao2008,assefa2009}. Genomic recombination
can however reduce the efficacy of this. \pb

Repeat regions may also be responsible for multiple gaps when building a
genome sequence; tandemly repeated nucleotide regions in the genome produce
multiple reads with similar sequence. As many assembly algorithms rely on
sequence overlaps between reads to build a contig, the similarity between
repeat-region reads can result in the assembly collapsing into an artificially
short sequence or being ignored by more conservative assembly algorithms. Such
regions can be resolved by using algorithms that specifically reassemble the
collapsed repeat region \cite{mulyukov2002,koren2010}. A related approach uses
unassembled sequence reads matching the regions around a scaffold gap to
construct a uniquely overlapping set of reads across the gap. \cite{tsai2010}.
\pb

Paired-read data can provide an extra level of information about how contigs
may be scaffolded together. Heuristic scaffolding algorithms take advantage of
this data to search for the optimal configuration of contigs in the scaffold
that matches these paired-read distances \cite{dayarian2010,boetzer2011}.
Synteny data from a reference genome can also be combined with this paired-read
data to estimate the best contig configuration \cite{pop2004}. \pb

These described \emph{in silico} methods provide a wide array of approaches for
merging contigs into a larger, continuous scaffold sequence. The scaffolding
process may still require manually inserting additional sequences or further
joining contigs using PCR-derived sequence. Moving and editing large blocks of
nucleotide text by hand however introduces human error and precludes any
reproducibility. \pb

The software outlined here, ``Scaffolder,'' aims to address these problems of
reproducibility by creating a file syntax and software framework for editing a
genome scaffold. Scaffolder uses a specific file format to define how contigs
are joined, additional sequences are inserted, and for the specification of
unknown regions. This syntax allows a scaffold to be updated by simply editing
the scaffold file. As such, Scaffolder facilitates a reproducible finishing
process and provides a concise overview of how the final genomic scaffold was
constructed. \pb

\clearpage

\section*{Implementation} %{{{1

\subsection*{Code and Dependencies} %{{{2

Scaffolder is written in the Ruby programming language and tested against
versions 1.8.7 and 1.9.2 \cite{ruby-lang}. The Scaffolder package is split into
two libraries. The first called ``scaffolder'' which provides the core
Scaffolder application programming interface (API). The second library
``scaffolder-tools'' provides the Scaffolder command line interface (CLI). \pb

Unit tests were implemented to maintain individual elements of the source code
during development and were written using the Shoulda and RSpec \cite{rspec}
libraries. Integration tests were written to test the Scaffolder software
interface as a whole and were written using the Cucumber library \cite{rspec}.
\pb

The Scaffolder source code is documented using the Yard library \cite{yard}.
Unix manual pages for the command line were generated using the Ronn library
\cite{ronn}. The manipulation of biological sequences in Scaffolder uses the
BioRuby library \cite{goto2010}. A full list of the software dependencies in
Scaffolder can be found in the Gemfile in the root of each source code
directory. \pb

\subsection*{Scaffold File Syntax} %{{{2

The choice of nucleotide sequences comprising the scaffold is specified using
the YAML syntax \cite{yaml}. YAML is a data format using whitespace and
indentation to produce a machine readable structure. As YAML is a standardised
data format, third-party developers have the option to generate a genome
scaffold using any programming language for which a YAML library exists. The
YAML website lists current parsers for languages including C/C++, Ruby, Python,
Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being widely
supported, YAML-formatted scaffold files can be validated for correct syntax
using third-party tools such as Kwalify \cite{kwalify}. \pb

Initial sequencing data assembly may result in an incomplete genome build.
Adding further sequences from either PCR or computational methods also means
that genome scaffolding may be an on-going process. The scaffold file should
therefore be simple to update manually in addition to being computationally
tractable. This requirement was also best suited to YAML syntax which is
human-readable and simple to edit in a standard text editor. \pb

The scaffold file takes the form of a list of entries. Each entry corresponds
to a region of sequence used in the final scaffold sequence. Each entry in the
scaffold file may have attributes that define whether a sub-sequence or the
reverse complement of the sequence should be used. The types of attributes
available, and an example scaffold file are outlined in the Results section.
\pb

The input data for Scaffolder is nucleotide sequence in FASTA format file.
These nucleotide sequences can be of any length and for example may be
individual reads, assembled contigs or contigs which have been joined into
larger scaffolds. The case in which Scaffolder may be most useful is using the
contigs and scaffolded contigs, combined with additional gap filling sequences
produced by PCR or \emph{in silico} methods as outlined in the Background. \pb

\clearpage

\section*{Results} %{{{1

\subsection*{Scaffolder Simplifies Genome Finishing} %{{{2

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in YAML specify how these sequences should be joined.
The Scaffolder software is used to generate the scaffold sequence from these
instructions. In addition to specifying which contigs are required, the
scaffold file allows the contigs to be edited into smaller sub-sequences or
reverse complemented if necessary. Each scaffold file represents one scaffolded
nucleotide sequence and separate scaffolds should be defined in separate files.
\pb

The process of genome finishing may involve producing additional
oligonucleotide sequences to fill unknown regions in a scaffold. The Scaffolder
format provides functionality to use these additional insert sequences to fill
gaps. These inserts can also be treated in the same manner as larger contig
sequences: trimmed and/or reverse complemented to match the corresponding gap
region size and orientation. \pb

The distances between contigs may be estimated from paired-read data or from
mapping the contigs to a reference genome. These inter-contig gap regions are
useful to join separate sequences together by the estimated distance. The
scaffold file allows for the specification of such unresolved regions by
inserting regions of `N' nucleotides into the scaffold. The use of these
regions in the scaffold indicates the unresolved regions in the build and their
approximate size. \pb

The nucleotide sequences used in the scaffold are maintained as a separate
FASTA file. The nucleotide sequences are referenced in the scaffold using the
first word from the FASTA header of the corresponding sequence. Maintaining the
nucleotide sequences in a separate file preserves the unedited sequence and
decouples the data from the specification of how it should be used to produce
the genome sequence. \pb

\subsection*{Defining a Scaffold as a Text File} %{{{2

The scaffold file is written using the YAML syntax and an example is shown in
Figure 1. This file illustrates the text attributes used to describe a scaffold
and how the sequences are correspondingly joined together in the genome build.
The basic layout of the scaffold file is a list of entries, where each entry
corresponds to a region of sequence in the generated scaffold super-sequence.
\pb

\subsubsection*{Simple sequence region} %{{{3

The first line of the scaffold file in Figure 1 begins with three dashes to
indicate the start of a YAML-formatted document. The first entry (highlighted
in green) begins with a dash character `-' to denote an entry in the YAML list.
This is a requirement of the YAML format: each entry begins with a dash line.
The next line is indented by two spaces where whitespace is used to group
similar attributes together. The ``sequence'' tag indicates that this entry
corresponds to a sequence in the FASTA file and the following line indicates
the name of this sequence using the ``source'' tag. The first word of the FASTA
header is used to identify which sequence is selected from the file. Together
these three lines describe the first entry in the scaffold as a simple sequence
using a FASTA entry named `sequence1'. On the right hand side of Figure 1 this
produces the first region in the scaffold, also shown in green. \pb

\subsubsection*{Unresolved sequence region} %{{{3

The second entry in the scaffold, highlighted in orange, is identified by the
``unresolved'' tag, indicating a region of unknown sequence but known length.
The second line specifies the size of this unknown region. This entry produces
a region of 20 `N' characters in the scaffold. \pb

\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3

The last entry in the scaffold, highlighted in blue, adds a FASTA entry named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags trim
the sequence to these coordinates inclusively. The ``reverse'' tag also
instructs Scaffolder to reverse complement the sequence. In the putative
scaffold shown in Figure 1 this completes the sequence. \pb

This final entry in the scaffold uses the ``inserts'' tag to add additional
regions of sequence. These inserts are also added as a YAML list, with each
insert beginning with a dash. The first insert, shown in purple, uses similar
attributes to that of a sequence entry; the reverse, start and stop tags are
used to trim and reverse complement the insert. Similarly the `source' tag
identifies the corresponding FASTA sequence as `insert1'. The ``open'' and
``close'' tags are specific to inserts and determine where the insert is added
in the enclosing sequence. The region of the sequence inside these coordinates
is inclusively replaced by the specified insert sequence. This is visualised in
the putative scaffold in Figure 1 by the black lines bisecting the blue
sequence. \pb

The next insert, shown in brown, is specified using only the `open' tag. This
illustrates that only one of either the `open' or `close' tags is required when
adding an insert sequence. If only one of the `open' or `close' tags is used
the corresponding opposing `open'/`close' coordinate is calculated from the
length of the insert FASTA sequence. This allows inserts to bridge into, and
partially fill gap regions, without requiring an end coordinate position. \pb

\subsection*{Scaffolder Software Interface} %{{{2

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) to validate the
scaffold file and build the draft super sequence. The Scaffolder CLI behaves as
a standard Unix tool and returns appropriate exit codes and manual pages. The
use of both these Scaffolder interfaces is outlined in detail on the Scaffolder
website (\scaffolder). This website provides a ``getting started'' guide as an
introduction to using Scaffolder to build a genome scaffold. \pb

\clearpage

\section*{Discussion} %{{{1

Scaffolding an incomplete genome assembly requires joining contigs and
additional gap-filling sequences using a combination of computational and
laboratory methods. The process of manually editing a scaffold is inherently
hard to reproduce and introduces irreproducible edits and/or human error. In
respect to this the aims of the Scaffolder software are twofold: 1) to provide
software that is easy to install and simplifies the task of genome finishing;
and 2) to facilitate reproducibility in the scaffolding and finishing stage of
a genome project. Scaffolder uses a minimal and compact syntax to describe how
a genome scaffold sequence should be generated. This syntax is simple to write
and edit whilst being succinct and readable. \pb

AGP is a similar format for describing scaffolds. This format can be used to
describe contig order and N-filled gap regions in a scaffold. The advantage of
the AGP format is that each contig entry is defined on a single line which
allows searching the scaffold using UNIX line-based tools. The Scaffolder
format in contrast is written in the standardised YAML format and therefore
accessible to the many languages which provide parsers to this format. The
Scaffolder format is provided with a tool explicitly to produce the FASTA
sequence of scaffold specified by the file. The Scaffolder format further
provides functions for trimming and replacing regions of sequence using
inserts. \pb

Constructing a genome by specifying the scaffold organisation in text file
makes generating a scaffold super sequence both reproducible and deterministic
for the same file and set of FASTA sequences. In comparison, joining large
nucleotide sequences by hand cannot be reliably reproduced. The scaffold file
furthermore provides a human readable description of how the scaffold is
constructed. Configuring the final sequence in the scaffold file means the
build is easier to edit once already constructed. \pb

A use case for Scaffolder is a combination of computational and manual editing
of a genome scaffold. We have used Scaffolder in our own genome projects to
create an initial scaffold from computationally parsing the output of \emph{in
silico} scaffolding tools into YAML. This scaffold was then manually updated as
the scaffold was finished with additional gap-filling sequences generated in
the laboratory. This is example of the scaffolder format being both
computationally tractable while being simple to edit manually. \pb

The YAML text format also allows comparison of differences between scaffold
builds using standard Unix tools such as diff. This therefore makes scaffold
files amenable to storage in version control systems and allows genome
finishers to use methods similar to those in software development. \pb

\clearpage

\section*{Conclusions} %{{{1

Scaffolder is software aimed at both bioinformaticians and biologists familiar
with the command line who wish to build a genome scaffold from a set of
contigs. The Scaffolder file format maintains the genome scaffold as a concise
and readable text representation that allows third-parties to see how the
genome sequence was scaffolded. This file format also allows a broad overview
of which sequences were included and how they are ordered in the genome
scaffold, something not possible to deduce from a megabase-length string of
nucleotide characters. Scaffolder furthers increases the ease of
reproducibility in genome projects by allowing the scaffold super-sequence to
be reliably reproduced from the same scaffold file. The YAML syntax for writing
the scaffold file is also standardised and simple to manipulate
programmatically. This thereby means the scaffolding process follows the Unix
tenet of ``If your data structures are good enough, the algorithm to manipulate
them should be trivial.''

\clearpage

\section*{Availability and Requirements} %{{{1

  \begin{description}
    \item[Project name:] Scaffolder v0.4.4, Scaffolder Tools
    v0.1.3
    \item[Project home page:] \scaffolder
    \item[Operating system:] Platform Independent. Tested on Mac OS X and
    Ubuntu.
    \item[Programming language:] Ruby 1.8.7 or 1.9.2
    \item[Other requirements:] RubyGems package management
    software and the following libraries: BioRuby 1.4.x, confligliere
    0.1.x, ronn 0.7.x. A full list of development dependencies can be found in
    the Gemfile in the base directory of each project.
    \item[License:] MIT \item[Any restrictions to use by non-academics:] None
    \end{description}

\clearpage

\section*{Competing Interests} %{{{1

The authors declare no competing interests.

\section*{List of Abbreviations Used} %{{{1

API: Application programming interface, CLI: Command line interface, PCR:
Polymerase chain reaction, YAML: YAML ain't markup language\cite{yaml}.

\section*{Authors contributions} %{{{1

MDB developed and maintains the Scaffolder tool. MDB and HAB wrote the
manuscript. All authors have read and approved the manuscript.

\section*{Acknowledgements} %{{{1
  \ifthenelse{\boolean{publ}}{\small}{}

This work was supported by the National Institute for Health: IDeA Network of
Biomedical Research Excellence (KY-INBRE) grant (NIH 2P20 RR016481-09) and the
NIH R15 AREA Program grant (R15GM079775).

\clearpage

% Bibliography {{{1
{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{article} }     % Bibliography file (usually '*.bib' )

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

\clearpage

\section*{Figures} %{{{1

\subsection*{Figure 1 - Example of Scaffolder File and the Resulting Build}

An example scaffold file written using the YAML syntax \cite{yaml} (left) and
the resulting putative scaffold sequence (right). The scaffold contains three
entries and two inserts. Each entry in the scaffold file text is delimited by a
`-' on a new line and highlighted using separate colours. The scaffold diagram
on the right is not to scale and instead illustrates how the scaffold sequences
are joined. \pb

\end{bmcformat}

\end{document}